---
title: "Setting up input files"
author: "Frederick Boehm"
date: "`r Sys.Date()`"
output: bookdown::html_document2
---

Our goal is to prepare the directory structures and files needed for input to the C++ compiled program, cvplus.

## What are the input arguments (for pheno1 in the UKB continuous data analysis)?

--n_fold 5 

--dbslmm_output_file_prefix ~/research/ukb-intervals/05_internal_c/pheno1/DBSLMM/summary_ukb_cross

--plink_file_prefix /net/mulan/disk2/yasheng/predictionProject/plink_file/ukb/chr

--alpha 0.1

--path_to_indicator_files ~/research/ukb-intervals/03_subsample/continuous/pheno1/indicator_files/

--path_to_true_pheno_file ~/research/ukb-intervals/03_subsample/continuous/pheno1/true_pheno.txt

--outpath ~/research/ukb-intervals/05_internal_c/pheno1/cvplus_outputs/

## Which inputs require new files? 

The "indicator files" and the "true pheno file" are the only new files needed. We'll create them with R code from the existing index files.

We already have the DBSLMM output files on the file server. Sheng previously created the plink files that are needed.

### Creating the indicator files from the existing subject ID files

We currently have files that record membership in the verification set and the fold-specific training and test sets. They use 
subject IDs. Subject IDs were created by Sheng. 

We need to read the pheno1 fam file to get the full list of subject IDs: /net/mulan/disk2/yasheng/predictionProject/plink_file/ukb/chr1.fam

We have the verification set IDs in the file:  ~/research/ukb-intervals/03_subsample/continuous/pheno1/verif/ukb/01_idx.txt

```{r}
library(magrittr)
pheno_number <- 1

fam_file <- "/net/mulan/disk2/yasheng/predictionProject/plink_file/ukb/chr1.fam"
fam <- readr::read_table(fam_file, col_names = FALSE)
training_indicator <- list()
for (i in 1:5){
    training_ids <- fam %>%
                        dplyr::filter(!is.na(.data[[paste0("X", pheno_number * 5 + 1)]]) | 
                                        !is.na(.data[[paste0("X", pheno_number * 5 + 2)]]) | 
                                        !is.na(.data[[paste0("X", pheno_number * 5 + 3)]]) | 
                                        !is.na(.data[[paste0("X", pheno_number * 5 + 4)]]) | 
                                        !is.na(.data[[paste0("X", pheno_number * 5 + 5)]])) %>%
                        dplyr::filter(is.na(.data[[paste0("X", pheno_number * 5 + 1)]]) | 
                                        is.na(.data[[paste0("X", pheno_number * 5 + 2)]]) | 
                                        is.na(.data[[paste0("X", pheno_number * 5 + 3)]]) | 
                                        is.na(.data[[paste0("X", pheno_number * 5 + 4)]]) | 
                                        is.na(.data[[paste0("X", pheno_number * 5 + 5)]])) %>%
                        dplyr::filter(!is.na(.data[[paste0("X", pheno_number * 5 + i)]])) %>%
                        dplyr::select(X1) %>%
                        unlist()

    training_indicator[[i]] <- fam$X1 %in% training_ids
}
```

```{r}
test_indicator <- list()


# subjects in fam file with all NA for a given trait truly have NA for that trait
for (i in 1:5){
    test_ids <- fam %>%
        dplyr::filter(!is.na(.data[[paste0("X", pheno_number * 5 + 1)]]) | 
                        !is.na(.data[[paste0("X", pheno_number * 5 + 2)]]) | 
                        !is.na(.data[[paste0("X", pheno_number * 5 + 3)]]) | 
                        !is.na(.data[[paste0("X", pheno_number * 5 + 4)]]) | 
                        !is.na(.data[[paste0("X", pheno_number * 5 + 5)]])) %>%
        dplyr::filter(is.na(.data[[paste0("X", pheno_number * 5 + i)]])) %>%
        dplyr::select(X1) %>%
        unlist()


    test_indicator[[i]] <- fam$X1 %in% test_ids
}
```

```{r}
dirpath <- "~/research/ukb-intervals/03_subsample/continuous/pheno1/indicator_files/"
# write the test & training indicator files, using as.numeric to convert to 1s and 0s.
for (fold in 1:5){
    readr::write_csv(x = tibble::as_tibble(as.numeric(training_indicator[[fold]])), 
                    file = paste0(dirpath, "indicator_training_fold", 
                                  fold, ".txt"),
                    col_names = FALSE
                    )
    readr::write_csv(x = tibble::as_tibble(as.numeric(test_indicator[[fold]])), 
                    file = paste0(dirpath, "indicator_test_fold", fold, ".txt"),
                    col_names = FALSE
                    )
}
```

```{r}
# write the verification set indicator for this pheno, using the 01_idx.txt file
verif_ids <- readr::read_table("~/research/ukb-intervals/03_subsample/continuous/pheno1/verif/ukb/01_idx.txt", 
                              col_names = FALSE)
verif_indic <- fam$X1 %in% verif_ids$X1
readr::write_csv(x = tibble::as_tibble(as.numeric(verif_indic)), 
                file = paste0(dirpath, "indicator_verification.txt"), 
                col_names = FALSE)
```

```{r}
# checks for indicator vectors
(as.numeric(training_indicator[[1]]) + as.numeric(test_indicator[[1]]) + as.numeric(verif_indic)) %>% max
(as.numeric(training_indicator[[1]]) + as.numeric(test_indicator[[1]]) + as.numeric(verif_indic)) %>% min

```

## Make the true pheno file

We lastly need to make a "true pheno file". This file will contain the true trait values for all subjects, regardless of 
membership in "training + test", validation or verification sets.

First, we'll read the fam file and select the relevant five columns. 
Then, we'll read the verification subject IDs. With the verification 
set IDs, we'll check to see if their trait values are included in the fam file.
At the moment, I'm unsure of the answer to this question.

```{r}
fam %>%
    dplyr::select(X1:X5, paste0("X", pheno_number * 5:(pheno_number * 5 + 5))) %>%
    dplyr::mutate(truth = purrr::pmap_dbl(.l = list(X6, X7, X8, X9, X10),
        .f = function(a, b, c, d, e){
            x <- c(a, b, c, d, e)
            x[!is.na(x)][1]
        }
    )) %>%
    dplyr::select(truth) %>%
    readr::write_csv(file = "~/research/ukb-intervals/03_subsample/continuous/pheno1/true_pheno.txt", col_names = FALSE)
```

